---
title: "Afghanistan Case Study" 
output: github_document
---

```{r setup and data visualization preferences}
library(tidyverse)
library(readxl)
library(p8105.datasets)
library(hexbin)
library(patchwork)
library(leaflet)
library(lubridate)
library(xml2)
library(rvest)

devtools::install_github("benmarwick/wordcountaddin",  type = "source", dependencies = TRUE)
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

##Explain 

Why we chose Afghanistan... 
Which datasets are being used... 

Using the years 2008-2016 because.... and supply code 

Aidworker dataset - cleaning and tailoring
```{r basic cleaning and exploring}
#Kailey notes: from section 1 cleaning and filtered for afghanistan/year, incld. pivot longer

url = "https://aidworkersecurity.org/incidents/search"
aidworker_html = read_html(url)

aidworker_df = 
  aidworker_html %>% 
  html_nodes(css = "table") %>%  
  first() %>% 
  html_table() %>% 
  as_tibble()

afghan_aidworker_df =
  aidworker_df %>%
  janitor::clean_names() %>% 
  select(-source, -verified) %>% 
  rename(year = year_sort_descending) %>% 
  mutate(intl_org_affected = 
           case_when(
             un != 0 ~ "yes",
             ingo != 0 ~ "yes",
             icrc != 0 ~ "yes",
             ifrc != 0 ~ "yes",
             other != 0 ~ "yes",
             lngo_and_nrcs != 0 ~ "no"),
         intl_org_affected = as.factor(intl_org_affected)) %>% 
  relocate(id, month, day, year, country, intl_org_affected) %>% 
  filter(country == "Afghanistan") %>% 
  filter(year %in% c("2008", "2009", "2010", "2011", "2012", "2013", "2014", "2015", "2016")) %>% 
    pivot_longer(
    un:other,
    names_to = "org_type", 
    values_to = "number_orgs_affected"
  ) %>% 
  select(-number_orgs_affected)

## Kailey notes: this date code is not working for me! 

aidworker_df$date <- as.Date(with(afghan_aidworker_df, paste(year, month, day, sep = "-")), "%Y-%m-%d")


```

Internationals vs. nationals killed (Not in the to-do list but has an interesting spike at 2010 and 2013... check historical context?)

```{r}
#Kailey notes: got this to work but the merging fix is probably not the right way to go about this.... I realized that the issue is that when you merge similar dataset, it creates duplicates. There is a bunch of information on how to fix this on stackoverflow if you search 'merge is duplicating rows in r' but I could not figure out what they wanted me to do so i used my slice-ing lol. Realizing that slice is my best friend but likely not the 'right' way or simplest way to go about things.... if you have any ideas or want to try the stack overflow fix, go for it! The numbers I get using this code seem reasonable, however! 

afghan_aidworker_international_df =
  afghan_aidworker_df %>% 
  group_by(year) %>% 
  count(internationals_killed) %>% 
  mutate(
    internationals_killed_tot = n
  )

afghan_aidworker_national_df = 
afghan_aidworker_df %>% 
  group_by(year) %>%
  count(nationals_killed) %>% 
  mutate(
    nationals_killed_tot = n
  ) 

afghan_aidworker_df_new = 
  left_join(afghan_aidworker_international_df, afghan_aidworker_national_df, by = "year") %>% 
  distinct(year, internationals_killed_tot, nationals_killed_tot) %>% 
  slice(1,4,7,10,13,16,20,24,28,32,34,36,40,44,48,52,56,60) %>% 
  group_by(year) %>% 
  mutate(
    yearly_internationals_killed = sum(internationals_killed_tot),
    yearly_nationals_killed = sum(nationals_killed_tot)) %>% 
  select(year, yearly_internationals_killed, yearly_nationals_killed) %>% 
  slice(1,4,6,9,12,14,17,21,23)


afghan_aidworker_df_new %>% 
     ggplot(aes(x = year, y = yearly_internationals_killed, col = "International")) +
   geom_line() +
  geom_line(aes(x = year, y = yearly_nationals_killed, col = "National")) +
  labs(
    title = "Number of Aidworkers Killed By Origin (National vs. International)",
    x = "Year",
    y = "Number Killed"
  ) 
```

# Filter Uppsala dataset for just Afghanistan to examine overall violence overtime alongside violence against aidworkers 
# Deaths among all civilians vs. all aidworkers side by side
-- are these different questions? 

Uppsala data upload and cleaning/tailoring (downloaded data for Afghanistan only)

```{r}
#Looking at deaths among all civilians verses all aidworkers side by side

#Aidworkers: 
aidw_killed_df =   
afghan_aidw_df %>% 
  group_by(year) %>% 
  count(total_killed) %>%
  select(year, n) %>% 
  mutate(
    aidwk_killed = sum(n)) %>% 
  select(year, aidwk_killed) %>% 
  slice(1, 7, 11, 16, 21, 23, 30, 36, 43)


#All civilians:
afghan_upp_df = 
read_csv("./data/afghanistan_uppsala.csv") %>%
  janitor::clean_names() %>% 
  filter(year %in% c("2008", "2009", "2010", "2011", "2012", "2013", "2014", "2015", "2016")) %>% 
  select(year, deaths_civilians) %>% 
  group_by(year) %>% 
  count(deaths_civilians) %>% 
  mutate(
    civilian_deaths = sum(n)
  ) %>% 
  select(year, civilian_deaths) %>% 
  slice(1, 19, 39, 63, 84, 104, 123, 143, 164) 

##alisha, this slice function is creating doubles for 2009 onward... i tried slicing again and taking the first of each double and got the same thing returned... very confused... doesn't seem to create and issue for the plot but still might want to fix? 

afghan_upp_df %>% 
  ggplot(aes(x = year, y = civilian_deaths, col = "Civilian")) +
   geom_line() +
  geom_line(data = aidw_killed_df, aes(x = year, y = aidwk_killed, col = "Aidworker")) +
  labs(
    title = "Aidowrkers and Civilians Killed",
    x = "Year",
    y = "Number Killed"
  ) 
```

#SHCC Afghanistan dataset (need to bring in multiple years) 

```{r}
shcc2018_df = 
  read_excel("data/2018-shcc-attacks-data.xlsx", skip = 1) %>% 
  janitor::clean_names() %>% 
  rename_at(vars(matches("number_")), ~ str_remove(., "number_")) %>% 
  filter(country_name == "Afghanistan") %>% 
    mutate(tot_hc_workers_affected = 
           affected_healthworker_killed + 
           affected_healthworker_sgbv + 
           affected_healthworker_arrested + 
           affected_healthworker_injured + 
           affected_healthworker_kidnapped) %>% 
  mutate(tot_hc_facilities_affected = 
           impact_health_facility_num +
           event_health_facility_destroyed_num +
           event_health_facility_damaged_num +
           impact_health_facility_armed_entry_num +
           impact_health_supplies_taken_num +
           impact_health_facility_occupation_num
           ) %>% 
  mutate(tot_hc_transport_affected = 
          impact_health_transport_destroyed_num + 
           impact_health_transport_damaged_num +
           impact_health_transport_abducted_num) 

# these columns are just straight up full of N/As...........
shcc2018_df %>% 
  ggplot(aes(x = date_occurrence, y = affected_healthworker_killed)) +
   geom_line() +
  geom_line(data = shcc2018_df, aes(x = date_occurrence, y = event_health_facility_destroyed_num)) +
  labs(
    title = "Aidowrkers and HCFs Destroyed",
    x = "Date",
    y = "Number Killed"
  ) 

# If we want to do a bar graph comparing total HCWs affected vs. total healthcare facilities affected in 2018 and 2019? Something is very off here.......


shcc2019_df = 
  read_excel("data/2019-shcc-healthcare-afghanistan-data.xlsx", skip = 1) %>% 
  janitor::clean_names() %>%
  rename_at(vars(matches("number_")), ~ str_remove(., "number_")) %>% 
  filter(country_name == "Afghanistan") %>% 
  select(-"x12") %>%
  mutate(
    tot_hc_worker_affected = 
      affected_healthworker_killed + 
      affected_healthworker_kidnapped +
      affected_healthworker_threatened + 
      affected_healthworker_injured + 
      affected_healthworker_assaulted + 
      affected_healthworker_sgbv
    ) %>% 
  mutate(
      tot_hc_facilities_affected = 
        indicator_health_facility_destroyed_num +
        indicator_health_facility_damaged_num + 
        indicator_health_facility_armed_entry_num +
        indicator_health_supplies_taken_num
    ) %>% 
  mutate(
       tot_hc_transport_affected = 
         indicator_health_transport_destroyed_num +
         indicator_health_transport_damaged_num +
         indicator_health_transport_abducted_num
      )
```


# Discuss confounders (U.S. administration changes? Troop levels in country? Natural disastor?)