---
title: "Reading and tidying data"
author: E. Brennan Bollman
date: '`r format(Sys.time(), "%y-%m-%d")`'
output: github_document
---

```{r setup, INCLUDE = FALSE}
library(tidyverse)
library(rvest)
library(httr)
library(readxl)

knitr::opts_chunk$set(
  fig.width = 10,
  fig.asp = 0.6,
  out.width = "90%"
)

theme_set(theme_bw() + theme(legend.position = "bottom")) 

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d()
scale_fill_discrete = scale_fill_viridis_d()
```


## Aid Worker Security Database

URL: https://aidworkersecurity.org/incidents
Select: 'all' in all fields to obtain the most complete form of the data
Download csv

Note: not sure if we should try to scrape from web to make this more reproducible. 

### Read in data

```{r}
aidworker_df = 
  read_csv("data/security_incidents_2020-11-23.csv") %>% 
  janitor::clean_names()
```

I think this looks fairly tidy? Each incident_id is row. There are lots of variables, but I think mostly will be able to manipulate with `mutate` and `group_by` in further operations on dataset. Lmk what you all think. 

Data dictionary: https://aidworkersecurity.org/about

### Quick EDA

Most deadly countries for aid workers.

```{r}
aidworker_df %>%
  group_by(country) %>% 
  summarize(tot_affected_per_country = sum(total_affected, na.rm = TRUE)) %>% 
  mutate(rank = min_rank(desc(tot_affected_per_country))) %>% 
  filter(rank < 6) %>% 
  knitr::kable()
```

Means of attack

```{r}
aidworker_df %>%
  count(means_of_attack) %>%
  mutate(rank = min_rank(desc(n))) %>% 
  arrange(rank) %>% 
  knitr::kable()
```

International workers affected

```{r}
aidworker_df %>% 
  group_by(country) %>% 
  summarize(tot_intl_per_country = sum(total_internationals, na.rm = TRUE)) %>% 
  mutate(rank = min_rank(desc(tot_intl_per_country))) %>% 
  filter(rank < 6) %>% 
  knitr::kable()
```

National workers affected

```{r}
aidworker_df %>% 
  group_by(country) %>% 
  summarize(tot_natl_per_country = sum(total_nationals, na.rm = TRUE)) %>% 
  mutate(rank = min_rank(desc(tot_natl_per_country))) %>% 
  filter(rank < 6) %>% 
  knitr::kable()
```


Syria per year

```{r}
aidworker_df %>% 
  filter(country == "Syrian Arab Republic") %>% 
  group_by(year) %>% 
  summarize(syria_tot_affected = sum(total_affected, na.rm = TRUE)) %>% 
  mutate(rank = min_rank(desc(syria_tot_affected))) %>% 
  arrange(rank) %>% 
  knitr::kable()
```


```{r}
aidworker_df %>% 
  filter(country == "Afghanistan") %>% 
  group_by(year) %>% 
  summarize(syria_tot_affected = sum(total_affected, na.rm = TRUE)) %>% 
  mutate(rank = min_rank(desc(syria_tot_affected))) %>% 
  arrange(year) %>% 
  knitr::kable()
```


```{r}
aidworker_df %>% 
  pivot_longer(
    un:other,
    names_to = "org_type", 
    values_to = "number_orgs_affected"
  ) %>% 
  select(-number_orgs_affected)
```

```{r}
aidworker_df

aidworker_df %>% 
  ggplot(aes(x = year, y = total_affected, color = country)) + 
  geom_point()
```



## SHCC dataset - 'Safeguarding Health in Conflict Coalition'

URL: https://data.humdata.org/dataset/shcchealthcare-dataset
Downloaded Afghanistan and Syria 2019 data

### Afghanistan

```{r}
afghan_shcc_colnames = 
  read_excel("data/2019-shcc-healthcare-afghanistan-data.xlsx", n_max = 0) %>%
  names()

afghan_shcc_df = 
  read_excel("data/2019-shcc-healthcare-afghanistan-data.xlsx", skip = 2, col_names = afghan_shcc_colnames) %>% 
  janitor::clean_names()
```

### Syria

```{r}
syria_shcc_colnames = 
  read_excel("data/2019-shcc-healthcare-syria-data.xlsx", n_max = 0) %>%
  names()

syria_shcc_df = 
  read_excel("data/2019-shcc-healthcare-syria-data.xlsx", skip = 2, col_names = syria_shcc_colnames) %>% 
  janitor::clean_names()
```


```{r}
syria_shcc_df %>%
  group_by(country) %>% 
  mutate(tot_hcw_affected = 
           total_health_worker_killed + 
           total_health_worker_kidnapped + 
           total_health_worker_arrested + 
           total_health_worker_injured + 
           total_health_worker_assaulted) %>% 
  count(tot_hcw_affected)
```

## Upsala dataset

```{r}
ucdp_df = 
  GET("https://ucdpapi.pcr.uu.se/api/gedevents/20.1?pagesize=1&page=1") %>% 
  content("text") %>% 
  jsonlite::fromJSON() %>% 
  as_tibble()
  
```


```{r}
ucdp_onsided_df = 
  read_csv("data/ucdp-onesided-201.csv")
```

```{r}
ucdp_onsided_df %>% 
  filter(location == "Syria") %>% 
  select(year, actor_name, best_fatality_estimate, low_fatality_estimate, high_fatality_estimate, is_government_actor) %>% 
  arrange(year)
```


```{r}
ucdp_onsided_df %>% 
  filter(location == "Afghanistan") %>% 
  select(year, actor_name, best_fatality_estimate, low_fatality_estimate, high_fatality_estimate, is_government_actor) %>% 
  arrange(year)
```

Does not seem terribly useful?

## Armed Conflict Location & Event Data Project (ACLED)

Afghanistan data here: https://data.humdata.org/dataset/acled-data-for-afghanistan

```{r}
afghan_acled_df = 
  read_csv("data/conflict_data_afg.csv")
```


# Working on EDA for Section 1

### International vs National staff attacks

```{r}
aidworker_tidy_df %>%
  group_by(year) %>% 
  summarize(tot_national = sum(total_national_staff),
            tot_intl = sum(total_international_staff),
            tot_both = sum(total_victims),
            pct_intl = (tot_intl/tot_both)*100,
            pct_national = (tot_national/tot_both)*100) %>%
  knitr::kable()

## If look at last row of data, appears "NA" is the total

aidworker_tidy_df %>%
  group_by(year) %>% 
  summarize(tot_national = sum(total_national_staff),
            tot_intl = sum(total_international_staff),
            tot_both = sum(total_victims),
            pct_intl = (tot_intl/tot_both)*100,
            pct_national = (tot_national/tot_both)*100) %>% 
  ggplot(aes(x = year, y = pct_national)) + 
  geom_line()
```

Roughly this shows that proportions of nationals attacked has always been higher than internationals, with increasing percentage of national staff attacked over time. 

```{r}
aidworker_tidy_df %>% 
  group_by(year) %>% 
   summarize(tot_national = sum(total_national_staff),
            tot_intl = sum(total_international_staff),
            tot_both = sum(total_victims)) %>% 
  ggplot(aes(x = year)) + 
  geom_line(aes(y = tot_national, color = "National Staff")) + 
  geom_line(aes(y = tot_intl, color = "International Staff")) + 
  ylim(0, 500) + 
  labs(title = "Aid Worker Attacks over time",
       x = "Year",
       y = "Number of Aid Workers Attacked")
```


```{r}
aidworker_tidy_df %>% 
  group_by(year, means_of_attack) %>% 
  summarize(tot_national = sum(total_national_staff),
            tot_intl = sum(total_international_staff),
            tot_both = sum(total_victims))
```


#### Attack context and actor types

```{r}
aidworker_df %>% 
  drop_na(year) %>% 
  filter(attack_context != "Unknown") %>% 
  group_by(year, attack_context) %>% 
  summarize(tot_national = sum(total_national_staff),
            tot_intl = sum(total_international_staff),
            tot_both = sum(total_victims)) %>%
  ggplot(aes(x = year, y = tot_both, color = attack_context)) + 
  geom_point() + 
  labs(title = "Attack Contexts Over Time",
       x = "Year",
       y = "Aid Worker Victims")
```

This is interesting: ambush = attacks on roads are most common and have been increasing. 

Brennan: Is there a better way to visualize this?

```{r}
aidworker_df %>% 
  drop_na(year) %>%
  mutate(actor_type = recode(actor_type,
                             "Non-state armed group: National" = "Non-state armed group",
                             "Non-state armed group: Regional" = "Non-state armed group",
                             "Non-state armed group: Subnational" = "Non-state armed group",
                             "Non-state armed group: Unknown" = "Non-state armed group"),
         actor_type = recode(actor_type,
                             "State: unknown" = "State or paramilitary",
                             "Host state" = "State or paramilitary",
                             "Police or paramilitary" = "State or paramilitary")) %>%
  filter(actor_type == c("Non-state armed group", "State or paramilitary", "Foreign or coalition forces")) %>% 
  group_by(year, actor_type) %>% 
  summarize(tot_national = sum(total_national_staff),
            tot_intl = sum(total_international_staff),
            tot_both = sum(total_victims))
```

Tried to make actor types more interesting by only looking at State forces (including paramilitary), non-state forces, and foreign forces -- not sure it's very interesting
