---
title: "Exploratory Analyses"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide 
    theme: flatly
    always_allow_html: true
---

```{r include = FALSE}
library(tidyverse)
library(dplyr)
library(patchwork)
library(leaflet)
library(lubridate)
library(rvest)
library(httr)
library(flexdashboard)
library(plotly)
library(viridis)
library(stringr)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

```{r message = FALSE, warning = FALSE}
url = "https://aidworkersecurity.org/incidents/search"
aidworker_html = read_html(url)

aidworker_df = 
  aidworker_html %>% 
  html_nodes(css = "table") %>%  
  first() %>% 
  html_table() %>% 
  as_tibble()

aidworker_df =
  aidworker_df %>%
  janitor::clean_names() %>% 
  select(-source, -verified) %>% 
  rename(year = year_sort_descending) %>% 
  mutate(intl_org_affected = 
           case_when(
             un != 0 ~ "yes",
             ingo != 0 ~ "yes",
             icrc != 0 ~ "yes",
             ifrc != 0 ~ "yes",
             other != 0 ~ "yes",
             lngo_and_nrcs != 0 ~ "no"),
         intl_org_affected = as.factor(intl_org_affected)) %>% 
  mutate(
    latitude = as.numeric(latitude),
    longitude = as.numeric(longitude)
    ) %>% 
  relocate(id, month, day, year, country, intl_org_affected)

## This is a function that should turn strings with empty spaces into NA 
empty_as_na <- function(x){
    if("factor" %in% class(x)) x <- as.character(x) 
    ifelse(as.character(x)!="", x, NA)
}

aidworker_df = 
aidworker_df %>% mutate_each(funs(empty_as_na)) 

aidworker_df = 
  aidworker_df %>% 
  filter(country != "Total") %>% 
  mutate(attack_abr = case_when(
           means_of_attack == "Kidnap-killing" ~ "Kidnapping",
           means_of_attack == "Kidnapping" ~ "Kidnapping",
           means_of_attack == "Body-borne IED" ~ "IED", 
           means_of_attack == "Vehicle-born IED" ~ "IED", 
           means_of_attack == "Roadside IED" ~ "IED",
           means_of_attack == "Landmine" ~ "Explosives", 
           means_of_attack == "Shelling" ~ "Explosives",
           means_of_attack == "Other Explosives" ~ "Explosives",
           means_of_attack == "Aerial bombardment" ~ "Explosives",
           means_of_attack == "Rape/sexual assault" ~ "Rape/sexual assault",
           means_of_attack == "Complex attack" ~ "Complex attack",
           means_of_attack == "Shooting" ~ "Shooting",
           means_of_attack == "Unknown" ~ "Unknown", 
           means_of_attack == "Bodily assault" ~ "Bodily assault"
           ))
```

Modern humanitarian action began in 1859 with the founding of the Red Cross movement after activist Henri Dunant witnessed battlefield suffering in Solferino, Italy and envisioned a neutral party that could assist victims of conflict. Aidworkers today follow common core principles of humanity, neutrality, impartiality, and independence to assist any person in need: and as such, should be protected from violence themselves. Unfortunately, the world is becoming more dangerous for those who provide humanitarian assistance.

* [Founding of ICRC - 2min video](https://www.icrc.org/en/who-we-are/history/founding)



## Attack trends over time

```{r message = FALSE, warning = FALSE}
plot_1 = 
  aidworker_df %>% 
  group_by(year) %>% 
  summarize(tot_attacks = n_distinct(id),
            tot_national = sum(total_national_staff),
            tot_intl = sum(total_international_staff),
            tot_both = sum(total_victims)) %>% 
  mutate(text_label = 
           str_c("Total Victims: ", tot_both, "\nInternational Staff: ", tot_intl,
                 "\nNational Staff: ", tot_national)) %>% 
  plot_ly(
    x = ~year, y = ~tot_attacks, text = ~text_label,
          type = "scatter", mode = "markers")
  
layout(plot_1, title = "Aid Worker Attacks Over Time", xaxis = list(title = "Year"), yaxis = list(title = "Number of Attack Incidents"))
```





# Plots

## International vs National Staff

## Attack Types


## Means of Attack


